---
title: "Calculate neighborhood weights"
author: "Satya Golla, Mounika Jakkkidi and Steve Simon"
date: "Created 2022-03-11"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../results", output_format = "all") })
---
This program will calculate which census blocks intersect with which neighborhoods. It relies on information stored in


+ bg.RData
+ bl.RData
+ nbd.RData
+ nbd-intersections.RData
+ red.RData

### General information

For information about various files, refer to the files that stored the relevant data (e.g., store_block_information.Rmd).

It helps to know the FIPS (Federal Information Processing System) codes for Kansas (20) and Missouri (29).

The important counties for this program are

Cass County, MO (29037)
Clay County, MO (29047)
Jackson County, MO (29095)
Johnson County, KS (20091)
Leavenworth County, KS (20103)
Wyandotte County, KS (20209)

### Setup

```{r setup}
library(sf)
library(tidyverse)
path_name <- "../Data/"
```

### Load key files

Block groups

```{r load-bg}
load(paste0(path_name, "bg", ".RData"))
glimpse(bg)
```

Blocks

```{r load-bl}
load(paste0(path_name, "bl", ".RData"))
glimpse(bl)
```

neighborhoods

```{r load-nbd}
load(paste0(path_name, "nbd", ".RData"))
glimpse(nbd)
```

neighborhoods intersections

```{r load-nbd-intersections}
load(paste0(path_name, "nbd-intersections", ".RData"))
glimpse(bg_nbd_intersection)
glimpse(bl_nbd_intersection)
```

```{r load-redistricting}
load(paste0(path_name, "red", ".RData"))
glimpse(red)
```

### Counts

```{r display-nice-counts}
# This code will compute the count of a vector
# and display it with commas if it is greater
# than 999.
nice_count <- function(x) {
  x                                    %>%
    unique                             %>%
    length                             %>%
    format(big.mark=",")
}
```

There are 7 counties in the Kansas City metropolitan area.

There are `r nice_count(nbd$NID)` neighborhoods in the 7 counties.

There are 
`r nice_count(bg$bg_id)`
block groups in the 7 counties. 
`r nice_count(bg_nbd_intersection$bg_id)`
of these intersect partially or completely with one or more neighborhoods.

There are 
`r nice_count(bl$bl_id)`
blocks in the 7 counties. 
`r nice_count(bl_nbd_intersection$bl_id)` 
of these intersect partially or completely with one or more neighborhoods.

### Get block group list

Identify the list of all block groups that partially or completely intersect with any neighborhoods.

```{r intersect-bg}
bg_nbd_intersection                    %>%
  pull(bg_id)                          %>%
  unique                                -> bg_list
glimpse(bg_list)
```

### Get block list

Now get a list of all census blocks that lie inside any of these block groups. Merge with red to get population and housing unit counts.

```{r pull-blocks}
bl                                     %>%
  pull(bl_id)                          %>%
  str_sub(1, 12)                       %>%
  bind_cols(bl$bl_id)                  %>%
  set_names(c("bg_id", "bl_id"))       %>%
  
  filter(bg_id %in% bg_list)           %>%
  inner_join(red, 
    by=c("bg_id", "bl_id"))               -> bl_list
glimpse(bl_list)
```

### Calculate counts and areas of intersections

```{r calculate-proportions}
bl_nbd_intersection                     %>%
  full_join(bl_list, 
    by=c("bg_id", "bl_id"))            %>%
  replace_na(list(bl_prop_in=0))       %>%
  replace_na(list(NID=0))            %>%
  mutate(
    people_in=
      round(people*bl_prop_in))        %>%
  mutate(
    units_in=
      round(units*bl_prop_in))         %>%
  select(
    bg_id,
    bl_id,
    NID,
    bl_prop_in,
    people,
    people_in,
    units,
    units_in)                           -> bl_counts

glimpse(bl_counts)
```

### Calculate counts for block groups

```{r count-bl}
red                                    %>%
  group_by(bg_id)                      %>%
  summarize(
    people=sum(people),
    units=sum(units))                  -> count_total
glimpse(count_total)

bl_counts                              %>%
  group_by(bg_id, NID)                 %>%
  summarize(
    people_in=sum(people_in),
    units_in=sum(units_in))            %>%
  filter(NID > 0)                      -> count_in
glimpse(count_in)

count_in                               %>%
  inner_join(count_total, by="bg_id")  %>%
  filter(NID > 0)                      -> bg_counts
glimpse(bg_counts)
```
###sum across blockgroups to get tracts counts
```{r}
bg_counts     %>%
   mutate(tr_id=str_sub(bg_id,1,11))  %>%
  group_by(NID,tr_id)                 %>%
  summarize(
    people=sum(people),
    people_in=sum(people_in),
    units=sum(units),
    units_in=sum(units_in))           -> tr_counts
  
glimpse(tr_counts)
```


### Save the information in an RData file.

```{r save}
save(
  bg_counts,
  bg_list,
  bl_counts,
  bl_list,
  tr_counts,
  file="../Data/nbd-weights.RData")
```
