---
title: "Calculate intersection of blocks and neighorhoods"
author: "Satya Golla, Mounika Jakkidi, and Steve Simon"
date: "Created 2022-03-11"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../results", output_format = "all") })
---

This program will calculate which census blocks intersect with neighorhoods. It relies on information stored in

+ bg.RData
+ bl.RData
+ nbd.RData

### General information

For information about various files, refer to the files that stored the relevant data (e.g., store_block_information.Rmd).

It helps to know the FIPS (Federal Information Processing System) codes for Kansas (20) and Missouri (29).

The important counties for this program are

Cass County, MO (29037)
Clay County, MO (29047)
Jackson County, MO (29095)
Johnson County, KS (20091)
Leavenworth County, KS (20103)
Wyandotte County, KS (20209)

### Load relevant files

```{r setup}
library(sf)
library(tidyverse)
path_name <- "../Data/"
load(paste0(path_name, "bg", ".RData"))
load(paste0(path_name, "bl", ".RData"))
load(paste0(path_name, "nbd", ".RData"))
```

### Intersect blocks and neighorhoods

Identify the list of all blocks that partially or completely intersect with neighborhoods. Calculate the block area and add it back to the data frame.

```{r intersect-bl}
Sys.time()
bl                                     %>%
  st_intersection(nbd)                 -> bl_nbd_intersection
glimpse(bl_nbd_intersection)

bl_nbd_intersection                    %>%
  st_area                              %>%
  as.numeric                           %>%
  tibble                               %>%
  set_names("in_area")                 %>%
  bind_cols(bl_nbd_intersection)       %>%
  mutate(
    bg_id=str_sub(bl_id, 1, 12))       %>%
  mutate(
    bl_prop_in=in_area/bl_area)        %>%
  mutate(
    bl_prop_in=round(bl_prop_in, 3))   %>%   
  select(
    bg_id,
    bl_id, 
    AreaName,
    NID, 
    bl_prop_in)                       -> bl_nbd_intersection
Sys.time()
glimpse(bl_nbd_intersection)
```


### Intersect block groups and neighorhoods

Identify the list of all block groups that partially or completely intersect with neighborhoods.

```{r intersect-bg}
Sys.time()
bg                                     %>%
  st_intersection(nbd)                 -> bg_nbd_intersection
glimpse(bg_nbd_intersection)

bg_nbd_intersection                    %>%
  st_area                              %>%
  as.numeric                           %>%
  tibble                               %>%
  set_names("in_area")                 %>%
  bind_cols(bg_nbd_intersection)       %>%
  mutate(bg_prop_in=in_area/bg_area)   %>%
  mutate(
    bg_prop_in=round(bg_prop_in, 3))   %>%
  select(
    bg_id,
    NID,
    AreaName,
    bg_prop_in)                       -> bg_nbd_intersection
Sys.time()
glimpse(bg_nbd_intersection)
```

### Save the information in an RData file.

```{r save}
save(
  bl_nbd_intersection,
  bg_nbd_intersection,
  file="../Data/nbd-intersections.RData")
```
