---
title: "Download ACS data using tidycensus"
author: "Satya Golla, Mounika Reddy,and Steve Simon"
date: "Created 2022-09-04"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../results", output_format = "all") })  
---

This program downloads information from 
the American Community Survey and displays
it on a map.

### Load relevant files

```{r setup}
# Important abbreviations:
#    bg = block group
#    bl = block
#    cd = community district
#    co = county
#    nbd=neighborhood
#    tr = tract

library(glue)
library(magrittr)
library(sf)
library(tidyverse)
library(tidycensus)
library(dplyr)
path_name <- "../data/"
```

### Contents of bg

```{r}
load(glue("{path_name}bg.RData"))
glimpse(bg)
```

### Contents of bl

```{r}
load(glue("{path_name}bl.RData"))
glimpse(bl)
```

### Contents of cd

```{r}
load(glue("{path_name}cd.RData"))
glimpse(cd)
```

### Contents of co

```{r}
load(glue("{path_name}co.RData"))
glimpse(co)
```

### Subset co to seven counties

```{r}
clist <- c(
  "20091",
  "20103",
  "20209",
  "29037",
  "29047",
  "29095",
  "29165")
co                                     %>%
  filter(GEOID %in% clist)             -> co
glimpse(co)
```

### Get id values for all community districts

```{r}
cd                                     %>%
  tibble                               %>%
  distinct(cd_id)                      %>%
  arrange(cd_id)                       %>%
  pull(cd_id)                           -> cd_list
```

### Pull state information from ACS

The examples here are loosely based on
the [basic usage vignette][walk1] for tidycensus.

[walk1]: https://walker-data.com/tidycensus/articles/basic-usage.html

### Housing units--occupied and unoccupied

```{r}
# Note: Data Profiles (DP) are available in the five year
# ACS data only down to the tract level.
vlist<- c("DP04_0016", "DP04_0024", "DP04_0025", "DP04_0026")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

get_acs(
  geography = "state", 
  variables = vlist,
  year = 2020)

get_acs(
  geography = "county", 
  variables = vlist, 
  state="MO",
  year = 2020)

get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)
```

### B01001 tables
https://pearson.socialexplorer.com/data/ACS2016/metadata/?ds=SE&table=B01001

```{r}
# Note: Detail tables (B) are available in the five year
# ACS data down to the census block group level.
vlist <- c("B01001_003", "B01001_027")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)
```

### B13002 tables
https://api.census.gov/data/2020/acs/acs5/groups/B13002.html

```{r}
# But some estimates are still only available in the
# five year ACS survey at the tract level or higher.
vlist <- c("B13002_001", "B13002_002", "B13002_011")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)
```
###"B21001"tables

https://data.census.gov/cedsci/table?q=Cedar%20Rapids%20veterans&tid=ACSDT5Y2020.B21001

```{r}
vlist <- c("B21001_001", "B21001_002", "B21001_003")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)
```
###"B25001" tables
https://data.census.gov/cedsci/table?q=housing%20units&tid=ACSDT5Y2020.B25001

```{r}
vlist <- c("B25001_001", "B25002_001", "B25002_002", "B25002_003")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)
```
###"B25002"
https://data.census.gov/cedsci/table?q=Table%20B25002&tid=ACSDT1Y2019.B25002&hidePreview=true

```{r}
vlist <- c("B25002_002", "B25003_001", "B25003_002", "B25003_003")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)
```
###"B25124"
https://api.census.gov/data/2020/acs/acs5/groups/B25124.html

```{r}
vlist <- c("B25124_002", "B25124_003", "B25124_010", "B25124_017", "B25124_024", "B25125_031")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)
```


```{r}
vlist <- c("B25124_038", "B25124_039", "B25124_046", "B25124_053", "B25124_060", "B25124_067")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)
```
###"B28001"
https://data.census.gov/cedsci/all?q=B28001&d=ACS%201-Year%20Estimates%20Detailed%20Tables

```{r}
vlist <- c("B28001_001", "B28001_002", "B28001_011")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)
```
###"B25003"
https://api.census.gov/data/2020/acs/acs5/groups/B25003.html

```{r}
vlist <- c("B25003_001E", "B25003_002E", "B25003_003E")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)
```
###"B09001"
https://api.census.gov/data/2020/acs/acs5/groups/B09001.html

```{r}
vlist <- c("B09001_001", "B09001_002", "B09001_003","B09001_005")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)
```
###"B17001"
https://api.census.gov/data/2020/acs/acs5/groups/B17001.html

```{r}
vlist <- c("B17001_001", "B17001_002", "B17001_003","B17001_017","B17001_018")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)
```
###"B27001" tables
https://api.census.gov/data/2020/acs/acs5/groups/B27001.html

```{r}
vlist <- c("B27001_001", "B27001_002", "B27002_030")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)
```
###"B21001"tables

https://api.census.gov/data/2020/acs/acs5/groups/B992701.html

```{r}
vlist <- c("B992701_001", "B992701_002", "B992701_003")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)
```