---
title: "Map block groups and community districts"
author: "Steve Simon"
date: "Created 2023-04-06"
output: html_document
---

### Program purpose

This program will map a community district in green, add census blocks with red border, and place text inside the border.

### Load key functions

Running annotate-maps.R will create some useful functions.

+ conditional_glimpse
+ plot_green

```{r}
library(tidyverse)
run_tests <- TRUE # allow simple function tests
terse <- FALSE # display information on intermediate steps
source("annotate-maps.R") # load several functions
```

### Load key data

Key RData files are:

+ bg: shape files for census block groups
+ cd: shape files for community districts
+ cd-intersections: percentage area overlapping
+ cd-weights: estimated people/housing units in overlap

### bg

```{r}
path_name <- "../data/"
load(glue("{path_name}bg.RData"))
conditional_glimpse(bg)
```

### cd

```{r}
path_name <- "../data/"
load(glue("{path_name}cd.RData"))
conditional_glimpse(cd)
```

### cd-intersections

```{r}
path_name <- "../data/"
load(glue("{path_name}cd-intersections.RData"))
conditional_glimpse(bg_cd_intersection)
# cd-intersections.RData also contains
# information at the block and tract
# levels
```

### cd-weights

```{r}
path_name <- "../data/"
load(glue("{path_name}cd-weights.RData"))
conditional_glimpse(bg_counts)
# cd-weights.RData also contains bl_counts
```


```{r}
for(i_cd in cd$cd_id[1:2]) {
  cd0 <- filter(cd, cd_id==i_cd)
  bg0 <- find_bg(i_cd, lo=0.1)
  bg0$bg_area <- round(bg0$bg_area/(10^6), 2)
  bg_text <- align_tx(bg0, bg_counts, i_cd)
  ti <- glue("{cd0$cd_name} ({cd0$cd_id})")
  for (v in c("bg_area")) {
    ti <- glue("{v} for {cd0$cd_name} ({cd0$cd_id})")
    plot_green(cd0, bg0, bg0[[v]]) +
      ggtitle(ti)                        -> bg_plot
    plot(bg_plot)
  }
  for (v in c("people", "units")) {
    ti <- glue("{v} for {cd0$cd_name} ({cd0$cd_id})")
    plot_green(cd0, bg0, bg_text[[v]]) +
      ggtitle(ti)                        -> bg_plot
    plot(bg_plot)
  }
}
```
