---
title: "Map block groups and community districts"
author: "Steve Simon"
date: "Created 2023-04-06"
output: html_document
---

### Program purpose

This program will map a community district in green, add census blocks with red border, and place text inside the border.

### Load key functions

Running annotate-maps.R will create some useful functions.

+ conditional_glimpse
+ plot_green

```{r}
library(tidyverse)
run_tests <- TRUE # allow simple function tests
terse <- FALSE # display information on intermediate steps
source("annotate-maps.R") # load several functions
```

### Load key data

Key RData files are:

+ bg: shape files for census block groups
+ cd: shape files for community districts
+ cd-intersections: percentage area overlapping
+ cd-weights: estimated people/housing units in overlap

### bg

```{r}
path_name <- "../data/"
load(glue("{path_name}bg.RData"))
conditional_glimpse(bg)
```

### cd

```{r}
path_name <- "../data/"
load(glue("{path_name}cd.RData"))
conditional_glimpse(cd)
```

### cd-intersections

```{r}
path_name <- "../data/"
load(glue("{path_name}cd-intersections.RData"))
conditional_glimpse(bg_cd_intersection)
# cd-intersections.RData also contains
# information at the block and tract
# levels
```

### cd-weights

```{r}
path_name <- "../data/"
load(glue("{path_name}cd-weights.RData"))
conditional_glimpse(bg_counts)
# cd-weights.RData also contains bl_counts
```

```{r}
plot_cd_bg <- function(
    cd, 
    i_cd, 
    bg, 
    cross_list, 
    text, 
    lo=0, 
    hi=0) {
  # get common subsets
  cd0 <- filter(cd, cd_id==i_cd)
  cross_list                          %>%
    find_intersecting_bg(i_cd, lo, hi) -> bg_subset
  bg_counts                           %>%
    filter(bg_id %in% bg_subset)      %>%
    filter(cd_id == i_cd)             %>%
    arrange(bg_id)                    %>%
    pull(people)                       -> bg_text
  ti <- glue("{cd0$cd_name} ({cd0$cd_id})")
  plot_green(cd0, bg0, bg_text)        +
    ggtitle(ti))                       -> bg_plot
  plot(bg_plot)
}
```

```{r}
for(i_cd in cd$cd_id[1:2]) {
}
```