---
title: "Map block groups and community districts"
author: "Steve Simon"
date: "Created 2023-04-06"
output: html_document
---

### Program purpose

This program will map a community district in green, add census blocks with red border, and place text inside the border.

### Load key functions

Running annotate-maps.R will create some useful functions.

+ conditional_glimpse
+ plot_green

```{r}
library(tidyverse)
run_tests <- TRUE # allow simple function tests
terse <- FALSE # display information on intermediate steps
source("annotate-maps.R") # load several functions
```

### Load key data

Key RData files are:

+ bg: shape files for census block groups
+ cd: shape files for community districts
+ cd-intersections: percentage area overlapping
+ cd-weights: estimated people/housing units in overlap

### bg

```{r}
path_name <- "../data/"
load(glue("{path_name}bg.RData"))
conditional_glimpse(bg)
```

### cd

```{r}
path_name <- "../data/"
load(glue("{path_name}cd.RData"))
conditional_glimpse(cd)
```

### cd-intersections

```{r}
path_name <- "../data/"
load(glue("{path_name}cd-intersections.RData"))
conditional_glimpse(bg_cd_intersection)
# cd-intersections.RData also contains
# information at the block and tract
# levels
```

### cd-weights

```{r}
path_name <- "../data/"
load(glue("{path_name}cd-weights.RData"))
# merge with bg_cd_intersection to get
# bg_prop_in
bg_counts                             %>%
  inner_join(
    bg_cd_intersection, 
    by=c("bg_id", "cd_id"))           %>%
  mutate(
    bg_prop_in=round(100*bg_prop_in)) %>%
  mutate(
    short_id=str_sub(bg_id, 6, 12))    -> bg_counts
conditional_glimpse(bg_counts)
# cd-weights.RData also contains bl_counts

```

### Select interesting cds

```{r}
name_list <- c(
  "Brookside",
  "Midtown",
  "Swope Park",
  "Waldo",
  "Ward Parkway",
  "Greater Downtown",
  "Antioch",
  "Little Blue Valley",
  "Rosedale",
  "Argentine")

cd                                     %>%
  filter(cd_name %in% name_list)       %>%
  pull(cd_id)                           -> interesting_cd
```

### Check out bg_prop_in

```{r}
for(i_cd in sort(cd$cd_id)) {
  # create subsets
  cd0 <- filter(cd, cd_id==i_cd)
  bg0 <- find_bg(i_cd)
  bg0$bg_area <- round(bg0$bg_area/(10^6), 2)
  bg_text <- align_tx(bg0, bg_counts, i_cd)
  bg0$short_id <- str_sub(bg0$bg_id, 6, 12)

  # loop for text variables found in bg_text
  for (v in c("bg_prop_in")) {
    ti <- glue("{v} for {cd0$cd_name} ({cd0$cd_id})")
    plot_green(cd0, bg0, bg_text[[v]]) +
      ggtitle(ti)                        -> bg_plot
    plot(bg_plot)
  }
}
```

### Get counts of people and units

Look only at interesting community districts and focus only on those block groups that have more than 5% of their area inside the community district.

```{r}
for(i_cd in interesting_cd) {
  # create subsets
  cd0 <- filter(cd, cd_id==i_cd)
  bg0 <- find_bg(i_cd, lo=0.05)
  bg0$bg_area <- round(bg0$bg_area/(10^6), 2)
  bg_text <- align_tx(bg0, bg_counts, i_cd)
  bg0$short_id <- str_sub(bg0$bg_id, 6, 12)

  # loop for text variables found in bg_text
  for (v in c(
      "short_id", 
      "people", 
      "units")) {
    ti <- glue("{v} for {cd0$cd_name} ({cd0$cd_id})")
    plot_green(cd0, bg0, bg_text[[v]]) +
      ggtitle(ti)                        -> bg_plot
    plot(bg_plot)
  }
}
```

