---
title: "Quality check 03, List blocks in block groups split by community districts"
author: "Satya Golla and Steve Simon"
date: "Created 2023-03-30"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../results", output_format = "all") })  
---

This program lists blocks within block groups that only partially intersect with community districts.

### Load relevant files

```{r setup}
# Important abbreviations:
#    bg = block group
#    bl = block
#    cd = community district
#    co = county
#    nbd=neighborhood
#    tr = tract

library(glue)
library(magrittr)
library(sf)
library(tidyverse)
library(dplyr)
path_name <- "../data/"
```

### Test acs download

```{r}
library(tidycensus)
county_list <- c(
  "Johnson",
  "Leavenworth",
  "Wyandotte",
  "Cass",
  "Clay",
  "Jackson",
  "Platte")
state_list <- rep(c("KS", "MO"), c(3,4))

vlist <- c("B01001_003", "B01001_027")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

acs_children <- NULL
for (i in 1:7) {
  acs1 <- get_acs(
    geography = "cbg", 
    variables = vlist, 
    state=state_list[i],
    county=county_list[i],
    year = 2020)
  acs_children <- rbind(acs_children, acs1)
}
glimpse(acs_children)
acs_children %<>% 
  group_by(GEOID) %>% 
  summarize(estimate=sum(estimate))
glimpse(acs_children)
```

### Contents of bg

```{r}
load(glue("{path_name}bg.RData"))
glimpse(bg)
```

### Contents of bl

```{r}
load(glue("{path_name}bl.RData"))
glimpse(bl)
```

### Contents of cd

```{r}
load(glue("{path_name}cd.RData"))
glimpse(cd)
```

### Contents of co

```{r}
load(glue("{path_name}co.RData"))
glimpse(co)
```

### Contents of cd-intersections

```{r}
load(glue("{path_name}cd-intersections.RData"))
glimpse(bg_cd_intersection)
glimpse(bl_cd_intersection)
```

### Contents of cd-weights

```{r}
load(glue("{path_name}cd-weights.RData"))
glimpse(bg_counts)
glimpse(bg_list)
glimpse(bl_counts)
glimpse(bl_list)
```

### Contents of red

```{r}
load(glue("{path_name}red.RData"))
glimpse(red)
```

### Subset co to seven counties

```{r}
clist <- c(
  "20091",
  "20103",
  "20209",
  "29037",
  "29047",
  "29095",
  "29165")
co                                     %>%
  filter(GEOID %in% clist)             -> co
glimpse(co)
```

### Get id values for all community districts

```{r}
cd                                     %>%
  tibble                               %>%
  distinct(cd_id)                      %>%
  arrange(cd_id)                       %>%
  pull(cd_id)                           -> cd_list
```

### Draw an overview map

```{r}
co_points <- st_point_on_surface(co)
co_coords <- as.data.frame(st_coordinates(co_points))
co_coords$NAME <- co$NAME

ggplot()+
  geom_sf(
    data=co, 
    color="darkred", 
    fill=NA)                            +
  geom_text(
    data=co_coords, 
    aes(X, Y, label=NAME))              +
  xlab("Longitude")                     +
  ylab("Latitude")                      -> overview_map
plot(overview_map)
```

### High level view

```{r pull-interesting-cd-ids, echo=FALSE}
name_list <- c(
  "Brookside",
  "Midtown",
  "Swope Park",
  "Waldo",
  "Ward Parkway",
  "Greater Downtown",
  "Antioch",
  "Little Blue Valley",
  "Rosedale",
  "Argentine")

cd                                     %>%
  filter(cd_name %in% name_list)       %>%
  pull(cd_id)                           -> interesting_cd
```

```{r pull-information, echo=FALSE}
pick_bg <- function(da, i_bg) {
  # Use this with
  #    bg
  #    bl
  da                                  %>%
    filter(bg_id %in% i_bg)           %>%
    return
}

pick_bg_cd <- function(da, i_bg, i_cd) {
  # Use this with
  #   bg_cd_intersection
  #   bg_counts
  #   bl_counts
  #   red
  da                                  %>%
    filter(bg_id %in% i_bg)           %>%
    filter(cd_id %in% i_cd)           %>%
    return
}

pick_bl <- function(da, i_bg) {
  # Use this with
  #    bl
  da                                  %>%
    filter(bg_id %in% i_bg)           %>%
    return
}

pick_bl_cd <- function(da, i_bl, i_cd) {
  # Use this with
  #   bl_cd_intersection
  #   bl_counts
  #   red
  da                                  %>%
    filter(bl_id %in% i_bl)           %>%
    filter(cd_id %in% i_cd)           %>%
    return
}

pick_cd <- function(da, i_cd) {
  # Use this with
  #    cd
  da                                  %>%
    filter(cd_id %in% i_cd)           %>%
    return
}

```

```{r plot-pieces, echo=FALSE}
terse <- TRUE
calc_cd0 <- function(cd, i_cd) {
  cd0 <- cd[cd$cd_id==i_cd, ]
  if (terse) return(cd0)
  cat(glue("i_cd = {i_cd}"))
  cat("cd0")
  glimpse(cd0)
  return(cd0)  
}

plot_co_cd <- function(overview_map, cd0) {
  # Draw community district in green atop
  # counties outlined in red
  ti <- cd0$cd_name
  overview_map                          +
    ggtitle(ti)                         +
    geom_sf(
      data=cd0,
      aes(),
      fill="lightgreen",
      color=NA)
}

calc_bg0 <- function(bg_cd_interestion, i_cd) {
  bg_cd_intersection                   %>%
    filter(cd_id==i_cd)                %>%
    filter(bg_prop_in > 0.10)          %>%
    pull(bg_id)                         -> bg_list
  bg                                   %>%
    filter(bg_id %in% bg_list)          -> bg0
  if (terse) return(bg0)
  cat("bg0")
  glimpse(bg0)
  return(bg0)
}


print_block_and_bg_data <- function(i_bg) {
  bl %>% pick_bg(i_bg)                -> bl0

  bl_counts                          %>%
    pick_bg_cd(i_bg, i_cd)           %>%
    mutate(
      prop_in = round(100*prop_in))  %>%
    data.frame                        -> view_bl_counts

  bg_cd_intersection                 %>%
    pick_bg_cd(i_bg, i_cd)           %>%
    mutate(
      bg_prop_in = 
        round(100*bg_prop_in))       %>%
    data.frame                        -> view_bl_cd

  red                                %>%
    pick_bg(i_bg)                    %>%
    data.frame                        -> view_red

  print("merged data")
  view_red                           %>%
    select(-bg_id)                   %>%
    full_join(
      view_bl_counts, 
      by="bl_id")                    %>%
    select(
      bl_id,
      prop_in,
      people,
      units,
      people_in,
      units_in)                      %>%
    arrange(
      desc(prop_in), 
      desc(people))                  %>%
    data.frame                        -> view_merge
  view_merge                         %>%
    print

  print("Summation for merged data")
  view_merge                         %>%
    summarize(
      people_in=sum(people_in, na.rm=TRUE),
      units_in=sum(units_in, na.rm=TRUE),
      people=sum(people, na.rm=TRUE),
      units=sum(units, na.rm=TRUE))  %>%
    print
    
  print("bg_counts")
  bg_counts                          %>%
    pick_bg_cd(i_bg, i_cd)           %>%
    data.frame                        -> view_bg_counts
  view_bg_counts                     %>%
    print

  return(view_merge)
}
```

```{r, echo=FALSE}
calc_sf_text <- function(sf0, sf_id) {
  sf0                           %>%
    st_point_on_surface               %>%
    # st_coordinates                    %>%
    bind_cols(sf0[ , sf_id])     -> sf_coords
  if (terse) return(sf_coords)
  cat("sf_coords")
  glimpse(sf_coords)
  return(sf_coords)
}

plot_green <- function(sf1, sf2, sf_text) {
  # draw sf1 in green
  ggplot(sf1)                           +
    xlab("Longitude")                   +
    ylab("Latitude")                    +
      geom_sf(
        aes(),
        fill="lightgreen",
        color=NA)                       -> m1
  if (is.null(sf2)) return(m1)
  
  # draw sf2 with red border
  m1                                    +
    geom_sf(
      data=sf2,
      aes(),
      fill=NA,
      color="darkred")                  -> m2

  # add text from sf_text
  m2                                    +
    geom_sf_text(
      data=sf2,
      aes(label=.data[[sf_text]]))       -> m3
  return(m3)
}
```

```{r}
i_cd <- interesting_cd[1]
cd %>% pick_cd(i_cd) -> cd0
plot_green(cd0, co, "NAME")      +
  ggtitle(cd0$cd_name)           -> o_map
plot(o_map)
```


```{r loop, warnings=FALSE, message=FALSE, echo=FALSE}
for (i_cd in interesting_cd[1:3]) {
  # add extra information to bl and bg
  bl$bg_id <- str_sub(bl$bl_id, 1, 12)
  bg$ab <- str_sub(bg$bg_id, 6, 12)
  
  # Pick subsets
  cd %>% pick_cd(i_cd) -> cd0
  bg_cd_intersection                  %>%
    pick_cd(i_cd)                     %>%
    filter(bg_prop_in > 0.1)          %>%
    select(bg_id, bg_prop_in)          -> bg_list
  bg                                  %>%
    filter(bg_id %in% bg_list$bg_id)   -> bg0

  # plot overview
  plot_green(cd0, co, "NAME")      +
    ggtitle(cd0$cd_name)           -> o_map
  plot(o_map)

  # plot intersecting block groups
  plot_green(cd0, bg0, "ab") +
    ggtitle(cd0$cd_name)           -> g_map
  plot(g_map)

  # remove interior block groups
  bg_cd_intersection                  %>%
    pick_cd(i_cd)                     %>%
    filter(bg_prop_in > 0.1)          %>%
    filter(bg_prop_in < 0.9)          %>%
    select(bg_id, bg_prop_in)          -> bg_list
  if (nrow(bg_list)==0) next
  bg                                  %>%
    filter(bg_id %in% bg_list$bg_id)   -> bg0
  for (i_bg in bg_list$bg_id) {
    merged_data <- print_block_and_bg_data(i_bg)
 
    # Create subset of bl
    bl                                %>%
      filter(bg_id==i_bg)             %>%
      mutate(
        ab=str_sub(bl_id, 13, 15))     -> bl0
    
    # Calculate intersections of individual blocks
    cd0                         %>%
      st_intersection(bl0)       -> in0
    print("in0")
    glimpse(in0)

    # plot blocks with abbreviated ids
    ti <- glue("Block ids for {i_bg}")
    plot_green(in0, bl0, "ab")         +
      ggtitle(ti)                      -> j_map
    plot(j_map)
 
    # plot blocks with people
    bl0                                %>%
      inner_join(merged_data)           -> bl0
    ti <- glue("People counts for {i_bg}")
    plot_green(in0, bl0, "people")      +
      ggtitle(ti)                       -> p_map 
    plot(p_map)
  }
}
```

<<<<<<< HEAD
### Test acs download

```{r}
library(tidycensus)
county_list <- c(
  "Johnson",
  "Leavenworth",
  "Wyandotte",
  "Cass",
  "Clay",
  "Jackson",
  "Platte")
state_list <- rep(c("KS", "MO"), c(3,4))

vlist <- c("B01001_003", "B01001_027")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

acs_children <- NULL
for (i in 1:7) {
  acs1 <- get_acs(
    geography = "cbg", 
    variables = vlist, 
    state=state_list[i],
    county=county_list[i],
    year = 2020)
  acs_children <- rbind(acs_children, acs1)
}
glimpse(acs_children)
acs_children %<>% 
  group_by(GEOID) %>% 
  summarize(estimate=sum(estimate))
glimpse(acs_children)
```

=======
>>>>>>> 8688701b42b908fae0a7a86c8ddce9fc9fa12093
