---
title: "Quality check 08, Associate ACS data with neighborhood weights"
author: "Satya Golla, Mounika Jakkidi and Steve Simon"
date: "Created 2022-11-10"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../results", output_format = "all") })  
---

This program merges ACS data with the neighborhood weights (stored in bg_counts data frame).

### Load relevant files

```{r setup}
# Important abbreviations:
#    bg = block group
#    bl = block
#    cd = community district
#    co = county
#    nbd=neighborhood
#    tr = tract

library(dplyr)
library(glue)
library(magrittr)
library(sf)
library(tidycensus)
library(tidyverse)
path_name <- "../data/"
```

### Contents of nbd-weights

```{r}
load(glue("{path_name}nbd-weights.RData"))
glimpse(bg_counts)
```

### Load counts for houses that are occupied/vacant

```{r}
vlist <- c("B25001_001", "B25002_001", "B25002_002", "B25002_003")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

b25002_jackson <- get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

b25002_wyandotte <- get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="KS",
  county="Wyandotte",
  year = 2020)

b25002_jackson %>%
  bind_rows(b25002_wyandotte) -> b25002_all_counties

glimpse(b25002_all_counties)

```

### Merge with bg_counts

```{r}
b25002_all_counties %>%
  inner_join(bg_counts, by=c("GEOID"="bg_id")) %>%
  arrange(NID, GEOID, variable) -> b25002_nbd
b25002_nbd
```

### Sum across neighborhoods

```{r}
b25002_nbd %>%
  mutate(unit_weight=units_in/units) %>%
  mutate(estimate=estimate*unit_weight) %>%
  group_by(NID, variable) %>%
  summarize(estimate=sum(estimate)) -> b25002_weighted
glimpse(b25002_weighted)
b25002_weighted[1:10, ]
```

### Load counts for houses that are owners occupied

```{r}
vlist <- c("B25124_002", "B25124_003", "B25124_010", "B25124_017", "B25124_024", "B25125_031")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

b25124_jackson <- get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

b25124_wyandotte <- get_acs(
  geography = "tract", 
  variables = vlist, 
  state="KS",
  county="Wyandotte",
  year = 2020)

b25124_jackson %>%
  bind_rows(b25124_wyandotte) -> b25124_all_counties

glimpse(b25124_all_counties)

```

### Merge with tr_counts

```{r}
b25124_all_counties %>%
  inner_join(tr_counts, by=c("GEOID"="tr_id")) %>%
  arrange(NID, GEOID, variable) -> b25124_nbd
b25124_nbd
```
### Sum across neighborhoods

```{r}
b25124_nbd %>%
  mutate(unit_weight=units_in/units) %>%
  mutate(estimate=estimate*unit_weight) %>%
  group_by(NID, variable) %>%
  summarize(estimate=sum(estimate)) -> b25124_weighted
glimpse(b25002_weighted)
b25124_weighted[1:10, ]
```
### Load counts for types of computers in household.

```{r}
vlist <- c("B28001_001", "B28001_002", "B28001_011")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

b28001_jackson <- get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

b28001_wyandotte <- get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="KS",
  county="Wyandotte",
  year = 2020)

b28001_jackson %>%
  bind_rows(b28001_wyandotte) -> b28001_all_counties

glimpse(b28001_all_counties)

```

### Merge with bg_counts

```{r}
b28001_all_counties %>%
  inner_join(bg_counts, by=c("GEOID"="bg_id")) %>%
  arrange(NID, GEOID, variable) -> b28001_nbd
b28001_nbd
```
### Sum across neighborhoods

```{r}
b28001_nbd %>%
  mutate(unit_weight=units_in/units) %>%
  mutate(estimate=estimate*unit_weight) %>%
  group_by(NID, variable) %>%
  summarize(estimate=sum(estimate)) -> b28001_weighted
glimpse(b28001_weighted)
b28001_weighted[1:10, ]
```
### Load counts for owners occuppied.

```{r}
vlist <- c("B25003_001", "B25003_002", "B25003_003")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

b25003_jackson <- get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

b25003_wyandotte <- get_acs(
  geography = "cbg", 
  variables = vlist, 
  state="KS",
  county="Wyandotte",
  year = 2020)

b25003_jackson %>%
  bind_rows(b28001_wyandotte) -> b25003_all_counties

glimpse(b25003_all_counties)

```

### Merge with bg_counts

```{r}
b25003_all_counties %>%
  inner_join(bg_counts, by=c("GEOID"="bg_id")) %>%
  arrange(NID, GEOID, variable) -> b25003_nbd
b25003_nbd
```
### Sum across neighborhoods

```{r}
b25003_nbd %>%
  mutate(unit_weight=units_in/units) %>%
  mutate(estimate=estimate*unit_weight) %>%
  group_by(NID, variable) %>%
  summarize(estimate=sum(estimate)) -> b25003_weighted
glimpse(b25003_weighted)
b25003_weighted[1:10, ]
```
### Load counts for population under 18years by Age.

```{r}
vlist <- c("B09001_001", "B09001_002", "B09001_003","B09001_005")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

b09001_jackson <- get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

b09001_wyandotte <- get_acs(
  geography = "tract", 
  variables = vlist, 
  state="KS",
  county="Wyandotte",
  year = 2020)

b09001_jackson %>%
  bind_rows(b09001_wyandotte) -> b09001_all_counties

glimpse(b09001_all_counties)

```

### Merge with tr_counts

```{r}
b09001_all_counties %>%
  inner_join(tr_counts, by=c("GEOID"="tr_id")) %>%
  arrange(NID, GEOID, variable) -> b09001_nbd
b09001_nbd
```
### Sum across neighborhoods

```{r}
b09001_nbd %>%
  mutate(unit_weight=units_in/units) %>%
  mutate(estimate=estimate*unit_weight) %>%
  group_by(NID, variable) %>%
  summarize(estimate=sum(estimate)) -> b09001_weighted
glimpse(b09001_weighted)
b09001_weighted[1:10, ]
```
### Load counts for poverty status in the past 12 months by sex by age.

```{r}
vlist <- c("B17001_001", "B17001_002", "B17001_003","B17001_017","B17001_018")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)

b17001_jackson <- get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

b17001_wyandotte <- get_acs(
  geography = "tract", 
  variables = vlist, 
  state="KS",
  county="Wyandotte",
  year = 2020)

b17001_jackson %>%
  bind_rows(b17001_wyandotte) -> b17001_all_counties

glimpse(b17001_all_counties)

```

### Merge with tr_counts

```{r}
b17001_all_counties %>%
  inner_join(tr_counts, by=c("GEOID"="tr_id")) %>%
  arrange(NID, GEOID, variable) -> b17001_nbd
b17001_nbd
```
### Sum across neighborhoods

```{r}
b17001_nbd %>%
  mutate(unit_weight=units_in/units) %>%
  mutate(estimate=estimate*unit_weight) %>%
  group_by(NID, variable) %>%
  summarize(estimate=sum(estimate)) -> b17001_weighted
glimpse(b17001_weighted)
b17001_weighted[1:10, ]
```
### Load counts for Health insurance coverage status by sex by age

```{r}
vlist <- c("B27001_001", "B27001_002", "B27001_030")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)


b27001_jackson <- get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

b27001_wyandotte <- get_acs(
  geography = "tract", 
  variables = vlist, 
  state="KS",
  county="Wyandotte",
  year = 2020)

b27001_jackson %>%
  bind_rows(b27001_wyandotte) -> b27001_all_counties

glimpse(b27001_all_counties)

```

### Merge with tr_counts

```{r}
b27001_all_counties %>%
  inner_join(tr_counts, by=c("GEOID"="tr_id")) %>%
  arrange(NID, GEOID, variable) -> b27001_nbd
b27001_nbd
```
### Sum across neighborhoods

```{r}
b27001_nbd %>%
  mutate(unit_weight=units_in/units) %>%
  mutate(estimate=estimate*unit_weight) %>%
  group_by(NID, variable) %>%
  summarize(estimate=sum(estimate)) -> b27001_weighted
glimpse(b27001_weighted)
b27001_weighted[1:10, ]
```
### Load counts for allocation of Health insurance coverage  

```{r}
vlist <- c("B992701_001", "B992701_002", "B992701_003")
load_variables(2020, "acs5", cache = TRUE) %>%
  filter(name %in% vlist)


b992701_jackson <- get_acs(
  geography = "tract", 
  variables = vlist, 
  state="MO",
  county="Jackson",
  year = 2020)

b992701_wyandotte <- get_acs(
  geography = "tract", 
  variables = vlist, 
  state="KS",
  county="Wyandotte",
  year = 2020)

b992701_jackson %>%
  bind_rows(b992701_wyandotte) -> b992701_all_counties

glimpse(b992701_all_counties)

```

### Merge with tr_counts

```{r}
b992701_all_counties %>%
  inner_join(tr_counts, by=c("GEOID"="tr_id")) %>%
  arrange(NID, GEOID, variable) -> b992701_nbd
b992701_nbd
```
### Sum across neighborhoods

```{r}
b992701_nbd %>%
  mutate(unit_weight=units_in/units) %>%
  mutate(estimate=estimate*unit_weight) %>%
  group_by(NID, variable) %>%
  summarize(estimate=sum(estimate)) -> b992701_weighted
glimpse(b992701_weighted)
b992701_weighted[1:10, ]
```