---
title: "Quality check 07, Block group intersections with neighborhoods"
author: "Satya Golla, Mounika Jakkidi, and Steve Simon"
date: "Created 2022-03-21"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../results", output_format = "all") })  
---

This program examines block groups that only partially intersect with community districts.

In the cd.RData file, Rosedale has cd_id=209, Swope Park has cd_id=110.

Interesting neighborhoods might be Terrace Lake Gardens, Colonial Square, Hidden Valley, Crossgates because each of these coincides with a single tract.

Ruskin Heights Eastwood Hills East have four tracts totally inside. Bannister Acres has two tracts totally inside.

Swope Parkway-Elmwood has three tracts, one totally inside and two partially inside. Vineyard is also similar. Glenhaven, East Swope Highlands has two tracts, one totally inside and one mostly outside.

### Load relevant files

```{r setup}
# Important abbreviations:
#    bg = block group
#    bl = block
#    cd = community district
#    co = county
#    nbd=neighborhood
#    tr = tract

library(glue)
library(magrittr)
library(sf)
library(tidyverse)
library(dplyr)
path_name <- "../data/"
```

### Contents of bg

```{r}
load(glue("{path_name}bg.RData"))
glimpse(bg)
```

### Contents of bl

```{r}
load(glue("{path_name}bl.RData"))
glimpse(bl)
```

### Contents of nbd

```{r}
load(glue("{path_name}nbd.RData"))
glimpse(nbd)
```

### Contents of co

```{r}
load(glue("{path_name}co.RData"))
glimpse(co)
```

### Contents of cd-intersections

```{r}
load(glue("{path_name}nbd-intersections.RData"))
# glimpse(tr_nbd_intersection)
glimpse(bg_nbd_intersection)
glimpse(bl_nbd_intersection)
```

### Contents of nbd-weights

```{r}
load(glue("{path_name}nbd-weights.RData"))
glimpse(bg_counts)
glimpse(bg_list)
glimpse(bl_counts)
glimpse(bl_list)
```

### Subset co to seven counties

```{r}
clist <- c(
  "20091",
  "20103",
  "20209",
  "29037",
  "29047",
  "29095",
  "29165")
co                                    %>%
  filter(GEOID %in% clist)             -> co
glimpse(co)
```

### Get id values for all neighborhoods

```{r}
nbd                                    %>%
  tibble                               %>%
  distinct(NID)                        %>%
  arrange(NID)                         %>%
  pull(NID)                             -> nbd_list
```

### Draw an overview map

```{r}
co_points <- st_point_on_surface(co)
co_coords <- as.data.frame(st_coordinates(co_points))
co_coords$NAME <- co$NAME

ggplot()+
  geom_sf(
    data=co, 
    color="darkred", 
    fill=NA)                            +
  geom_text(
    data=co_coords, 
    aes(X, Y, label=NAME))              +
  xlab("Longitude")                     +
  ylab("Latitude")                      -> overview_map
plot(overview_map)
```

### High level view

```{r loop}
nbd                                    %>%
  pull(NID)                             -> interesting_nbd
for (i_NID in interesting_nbd) {
  # Draw community district in green atop
  # counties outlined in red
  nbd_subset <- filter(nbd, NID==i_NID) 
  ti <- nbd_subset$AreaName
  overview_map                          +
    ggtitle(ti)                         +
    geom_sf(
      data=nbd_subset,
      aes(),
      fill="lightgreen",
      color=NA)                         -> map1
  plot(map1)

  # draw community districts in green with
  # block groups outlined in red.
  bg_nbd_intersection                  %>%
    filter(NID==i_NID)                 %>%
    filter(bg_prop_in > 0.01)          %>%
    pull(bg_id)                         -> bg_list
  bg                                   %>%
    filter(bg_id %in% bg_list)          -> bg_subset

  bg_points <- st_point_on_surface(bg_subset)
  bg_coords <- as.data.frame(st_coordinates(bg_points))
  bg_coords$NAME <- str_sub(bg_subset$bg_id, 6, 12)

  ggplot(nbd_subset)                    +
    ggtitle(ti)                         +
    xlab("Longitude")                   +
    ylab("Latitude")                    +
      geom_sf(
        data=nbd_subset,
        aes(),
        fill="lightgreen",
        color=NA)                       +
    geom_text(
      data=bg_coords, 
      aes(X, Y, label=NAME))            +
    geom_sf(
      data=bg_subset,
      aes(),
      fill=NA,
      color="darkred")                  -> map2
  plot(map2)

  # draw block groups and blocks outlined
  # in red atop community district in green.
  for (i_bg in bg_list) {
    next
    bg_nbd_intersection                %>%
      filter(NID==i_NID)               %>%
      filter(bg_id==i_bg)              %>%
      pull(bg_prop_in)                  -> p
    if (p < 0.1) next
    if (p > 0.9) next
    bl_subset <- filter(bl, str_sub(bl_id, 1, 12)==i_bg)
    ggplot(nbd_subset)                  +
      ggtitle(ti)                       +
      xlab("Longitude")                 +
      ylab("Latitude")                  +
    geom_sf(
      data=nbd_subset,
      aes(),
      fill="lightgreen",
      color=NA)                         +
    geom_sf(
      data=bl_subset,
      aes(),
      fill=NA,
      color="darkred")                  -> map3
    plot(map3)
  }
}
```
